pirouette: the error BEAST2 makes in inferring a phylogeny

### 2019-01-29 article.tex
Use '\&' instead of '&', added fig 2

Phylogenies are commonly used to explore evolutionary hypotheses.
The mode of diversification is one such hypothesis.
Phylogenies constructed from DNA alignments are commonly fit to
diversification models, or the other way around. For example,
[Etienne & Rosindell 2012] fit four bird phylogenies to the protracted
diversification model.

There are plenty of diversification models: constant-birth [Yule, 19..],
birth-death [Nee et al., 19..], time-dependent [?], diversity 
dependent [Etienne & ?, approx 2012], protracted 
birth-death [Rosindell & Etienne, 2011][Etienne & Rosindell, 2012],
multiple-birth death [Laudanno, Bilderbeek & Etienne, in preparation],
a rate shifting birth-death model [the SLS model. Laudanno & Etienne, in preparation]
punctuated equilibrium aka early burst [?Harmon],
dependent on a binary trait [BISSE model, ?Maddison], 
continuous trait [HISSE model],
concealed state [?MUSSE] or a combination of all [secSSE model, Leonel].

Diversification models are getting more complex over time. There are many
reasons for this. One reason is the increase in computing power,
which allows this heightened complexity possible.
Another reason is to explore a process that is not
yet captured by the other diversification models. An example of this
is the multiple-birth model [Laudanno, Bilderbeek & Etienne, in preparation], 
that is the first model to allow speciation to co-occur.

@@

\verb;pirouette; is an R package that performs such an experiment,
and is built on \verb;babette; [Bilderbeek & Etienne, 2018], 
which calls the popular Bayesian inference tool 
BEAST2 [X and Bouckaert et al., ????]. With \verb;pirouette;, one
can easily measure the error made by Bayesian inference in recovering
any given phylogeny, helping us to valorize new diversification models.

@@

The fourth step measures the inference error, using the phylogenies in the
Bayesian posterior. These phylogenies are compared to the given
phylogeny using an error statistic, which is the nLTT 
statistic \cite{janzen2015approximate} by default,
but also an custom error statistic is supplied, based
on the absolute difference in the gamma statistic [Pybus & Harvy]. 
Additionally, the user can specify the
proportion of posterior phylogenies to 
discard (i.e. the burn-in), throwing away the first 10%
of all phylogenies by default. This burn-in is used to discard
the part of the MCMC chain that has not yet converged to a
representative part of state space.

@@

>>>>

Phylogenies are commonly used to explore evolutionary hypotheses.
The mode of diversification is one such hypothesis.
Phylogenies constructed from DNA alignments are commonly fit to
diversification models, or the other way around. For example,
[Etienne \& Rosindell 2012] fit four bird phylogenies to the protracted
diversification model.

There are plenty of diversification models: constant-birth [Yule, 19..],
birth-death [Nee et al., 19..], time-dependent [?], diversity 
dependent [Etienne \& ?, approx 2012], protracted 
birth-death [Rosindell \& Etienne, 2011][Etienne \& Rosindell, 2012],
multiple-birth death [Laudanno, Bilderbeek \& Etienne, in preparation],
a rate shifting birth-death model [the SLS model. Laudanno \& Etienne, in preparation]
punctuated equilibrium aka early burst [?Harmon],
dependent on a binary trait [BISSE model, ?Maddison], 
continuous trait [HISSE model],
concealed state [?MUSSE] or a combination of all [secSSE model, Leonel].

Diversification models are getting more complex over time. There are many
reasons for this. One reason is the increase in computing power,
which allows this heightened complexity possible.
Another reason is to explore a process that is not
yet captured by the other diversification models. An example of this
is the multiple-birth model [Laudanno, Bilderbeek \& Etienne, in preparation], 
that is the first model to allow speciation to co-occur.

@@

\verb;pirouette; is an R package that performs such an experiment,
and is built on \verb;babette; [Bilderbeek \& Etienne, 2018], 
which calls the popular Bayesian inference tool 
BEAST2 [X and Bouckaert et al., ????]. With \verb;pirouette;, one
can easily measure the error made by Bayesian inference in recovering
any given phylogeny, helping us to valorize new diversification models.

@@

The fourth step measures the inference error, using the phylogenies in the
Bayesian posterior. These phylogenies are compared to the given
phylogeny using an error statistic, which is the nLTT 
statistic \cite{janzen2015approximate} by default,
but also an custom error statistic is supplied, based
on the absolute difference in the gamma statistic [Pybus \& Harvy]. 
Additionally, the user can specify the
proportion of posterior phylogenies to 
discard (i.e. the burn-in), throwing away the first 10%
of all phylogenies by default. This burn-in is used to discard
the part of the MCMC chain that has not yet converged to a
representative part of state space.

@@

\begin{figure}[h]
  \includegraphics[width=\textwidth]{figure_2.png}
\end{figure}
